---
title: "Script5 PDX Translation Efficiency based on CDS counts"
output: html_document
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
---

We follow the analysis described in https://currentprotocols.onlinelibrary.wiley.com/doi/full/10.1002/cpmb.108 where they introduce an interaction term into the statistical model of DESeq2 to identify differential translation efficiency genes. When combining RNA-seq and Ribo-seq from two conditions, the interaction term can be used to model condition (untreated/treated) and sequencing methodology (Ribo-seq/RNA-seq). This allows the identification of significant differences between conditions that are discordant between sequencing methodologies. They design a generalized linear model with three components: the condition (c), the sequencing type (s), and an interaction term containing both (c:s). The result is a ΔTE fold change and an associated false discovery rate (FDR) for significant changes of this fold change, which quantify the extent of translational regulation between conditions.

```{r , echo=F, message=F, warning=F}
#Load libraraies
library(tximport)
library(openxlsx)
library(dplyr)
library(GenomicFeatures)
library(DESeq2)
library(stringr)
library(pheatmap)
library(plyr)
library(AnnotationDbi)
library(apeglm)
library(ggplot2)
library(EnhancedVolcano)
library(clusterProfiler)
library(org.Hs.eg.db)
library(fgsea)
library(msigdbr)
library(GGally)
library(tidyverse)
library(ggplotify)
library(ggpubr)
library(GGally)
library(knitr)
library(kableExtra)
library(ggVennDiagram)
```

```{r}
knitr::opts_chunk$set(echo = F, message = F, error=TRUE)
```

```{r}
gtf.file = "/local/storage/maa7095/references/human/gencode_primary/MANE_ref/MANE_select_subset_GRCh38.v1.2.ensembl_genomic.gtf"
gtf.gr = rtracklayer::import(gtf.file) # creates a GRanges object
gtf.df = as.data.frame(gtf.gr)
genes = unique(gtf.df[ ,c("gene_id","gene_name", "transcript_id")])
```


```{r, echo=F, message=F, warning=F}
#Load the metadata and get the file name from there.
TE_metadata <- read.xlsx("/SSD/maa7095/RIBOseq/NO_UMI_ANALYSIS/downstream_analysis/Xenograft_process/DeltaTE/TE_metadata_xeno.xlsx")
kable(TE_metadata, row.names=FALSE,  padding = 0, longtable=TRUE) %>%  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```



```{r, echo=F, message=F, warning=F}
RIBO_counts <- read.csv("/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/CDS_RIBO_raw_counts.csv")

#keep only trascript in MANE_select*not clinical)
RIBO_counts <- RIBO_counts %>% filter(transcript %in% genes$transcript_id) %>% rename(transcript_id=transcript)

#rename tramscript by gene symbol
genes <- genes %>% filter(transcript_id %in% RIBO_counts$transcript_id)
RIBO_counts <- join(RIBO_counts, genes, type="inner")
#There are three dupliacetd values
RIBO_counts = RIBO_counts[!duplicated(RIBO_counts$gene_name),]
rownames(RIBO_counts) <- RIBO_counts$gene_name
```


```{r, echo=F, message=F, warning=F}
RIBO_counts <- RIBO_counts %>% dplyr::select(gene_name, contains("_mapped_"))
RIBO_counts$gene_name <- NULL
names(RIBO_counts) <- sapply(strsplit(names(RIBO_counts), "_Aligned.toTranscriptome.out"), "[", 1)
names(RIBO_counts) <- sapply(strsplit(names(RIBO_counts), "human_mapped_"), "[", 2)
names(RIBO_counts) <- paste0("RIBO_seq_",names(RIBO_counts))
```

```{r, echo=F, message=F, warning=F}
#Load bulkRNAseq gene counts based on CDS and from featurecounts.

RNA_counts <- read.table("/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/CDS_RNA_counts_not_strand_xeno.txt", header = T, sep = "\t")

RNA_counts <- RNA_counts %>% select(Geneid, contains("_mapped_"))

RNA_counts <- RNA_counts %>% rename(gene_id=Geneid)
RNA_counts <- join(RNA_counts,genes)
RNA_counts = RNA_counts[!duplicated(RNA_counts$gene_name),]
RNA_counts <- RNA_counts %>% dplyr::select(gene_name, contains("_mapped_"))
rownames(RNA_counts) <- RNA_counts$gene_name

RNA_counts$gene_name <- NULL
names(RNA_counts) <- sapply(strsplit(names(RNA_counts), "_Aligned.sortedByCoord.out.bam"), "[", 1)
names(RNA_counts) <- sapply(strsplit(names(RNA_counts), "human_mapped_"), "[", 2)

names(RNA_counts) <- paste0("RNA_seq_",names(RNA_counts))
```


```{r, echo=F, message=F, warning=F}
#Bind matrices and check for rows in order.
all(rownames(RIBO_counts)==rownames(RNA_counts))
RIBO_counts <- RIBO_counts[rownames(RNA_counts),]
all(rownames(RIBO_counts)==rownames(RNA_counts))
```

```{r, echo=F, message=F, warning=F}
TE_metadata$Data_type <- factor(TE_metadata$Data_type, levels = c("RNA_seq", "RIBO_seq"))
TE_metadata$Condition <- factor(TE_metadata$Condition, levels = c("Primary","Met"))
```


```{r, echo=F, message=F, warning=F}
#Bind matrices and check for rows in order.
all(rownames(RIBO_counts)== rownames(RNA_counts))
TE_counts <- cbind(RIBO_counts,RNA_counts)
rownames(TE_metadata) <- TE_metadata$Sample_name
all(rownames(TE_metadata) ==colnames(TE_counts))
TE_counts <- TE_counts[,rownames(TE_metadata)]
all(rownames(TE_metadata) ==colnames(TE_counts))
```

```{r}
rna_metadata <- TE_metadata %>% filter(Data_type=="RNA_seq")
ribo_metadata <- TE_metadata %>% filter(Data_type=="RIBO_seq")
```



```{r,  message=F, echo=F, warning=F}
#Start runnig DESEQ2.
TE_ddsMat = DESeqDataSetFromMatrix(
  countData=round(TE_counts),
  colData=TE_metadata,
  design=~ Condition+Data_type+Condition:Data_type)
```

```{r, echo=F, message=F}
keep <- rowSums(counts(TE_ddsMat) >= 10) >= 2
TE_ddsMat <- TE_ddsMat[keep,]
TE_ddsMat <- DESeq(TE_ddsMat)
TE_normalized_counts <- counts(TE_ddsMat,normalized=TRUE)
TE_raw_counts <- counts(TE_ddsMat,normalized=F)

write.csv(TE_normalized_counts, "/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/TE_normalized_counts.csv")
write.csv(TE_raw_counts, "/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/TE_raw_counts.csv")

```


## TE analysis {.tabset}

First comparison in which three components: the condition (c), the sequencing type (s), and an interaction term containing both (c:s) will be take into account. Condition2.SeqTypeRibo.seq means Changes in Ribo-seq levels in Condition2 vs Condition1 accounting for changes in RNA-seq levels in Condition2 vs Condition1.

### PCA all datasets.
.
```{r, echo=F, message=F, warning=F}
TE_ddsMat_Coll <- rlog( TE_ddsMat, blind=FALSE)
TE_ddsMat_Coll_mat <- assay(TE_ddsMat_Coll)
TE_ddsMat_Coll_cor <- cor(TE_ddsMat_Coll_mat)
plotPCA(TE_ddsMat_Coll, intgroup=c("Condition", "Data_type")) 
```

### PCA RNA samples only.
```{r}
rna_samples <-TE_metadata[which(TE_metadata$Data_type == "RNA_seq"),]$Sample_name 
RNA_ddsMat = DESeqDataSetFromMatrix(
  countData=round(TE_counts[,rna_samples]),
  colData=TE_metadata[which(TE_metadata$Data_type == "RNA_seq"),],
  design=~Condition)
keep <- rowSums(counts(RNA_ddsMat) >= 10) >= 2
RNA_ddsMat <- RNA_ddsMat[keep,]
RNA_ddsMat <- DESeq(RNA_ddsMat)
RNA_ddsMat_Coll <- rlog( RNA_ddsMat, blind=FALSE)
RNA_ddsMat_Coll_mat <- assay(RNA_ddsMat_Coll)
RNA_ddsMat_Coll_cor <- cor(RNA_ddsMat_Coll_mat)

plotPCA(RNA_ddsMat_Coll, intgroup="Condition") 
rm(RNA_ddsMat)
```


### PCA all samples after removing the RNAseq outliers.
```{r}
excluded_columns <- c("RNA_seq_EP_11", "RNA_seq_EP_13")
TE_ddsMat <- TE_ddsMat[, -which(colnames(TE_ddsMat) %in% excluded_columns)]
TE_ddsMat <- DESeq(TE_ddsMat)
TE_ddsMat_Coll <- rlog( TE_ddsMat, blind=FALSE)
TE_ddsMat_Coll_mat <- assay(TE_ddsMat_Coll)
TE_ddsMat_Coll_cor <- cor(TE_ddsMat_Coll_mat)
plotPCA(TE_ddsMat_Coll, intgroup=c("Condition", "Data_type")) 
```


```{r}
results_TE = as.data.frame(results(TE_ddsMat, name="ConditionMet.Data_typeRIBO_seq"))
```


```{r, echo=F, message=F, warning=F}
results_TE <- results_TE[order(results_TE[,6]),]
  
write.csv(results_TE, file="/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/DESeq2_TE.csv", row.names = T, quote = F)
```

### Volcano plot for TE taking into account changes between RNA-seq. 
```{r, echo=F, message=F, warning=F, warning=F}
EnhancedVolcano(results_TE,lab = rownames(results_TE), x = 'log2FoldChange',y = 'pvalue', title = "TE analysis", pCutoff = 1e-05, FCcutoff = 1)
```

### Heatmap showing expression of DE genes padj < 0.05 LFC 1
```{r, echo=F, message=F, warning=F}
#draw heatmap with the most disregulated genes
padj_genes <-na.omit(results_TE) %>% filter(padj<0.05 ,  log2FoldChange > 1 |  log2FoldChange < -1)
padj_genes2 <- rownames(padj_genes)
print(paste0(length(padj_genes2), " DE genes under padjusted < 0.05 and LFC 1 in TE data"))

#Get the transformed values and filter genes
subset_mat <- TE_ddsMat_Coll_mat[padj_genes2,]
```


```{r, echo=F, message=F, warning=F}
pheatmap(subset_mat, scale="row", show_rownames = F, show_colnames = T, cluster_rows = T, legend = T, main =  "Heatmap of DE genes found in TE analysis", annotation_col = as.data.frame(colData(TE_ddsMat))[, c("Data_type", "Condition")])
```

## {-}

## RNAseq Met vs Primary DE results.{.tabset}

Second comparison Met_RNA vs Primary_RNA.

```{r, echo=F, message=F, fig.width=10, fig.height=10}
#Create bulkRNAseq object.
rna_samples <-grep("RNA", colnames(TE_ddsMat), value = T)
RNA_ddsMat = DESeqDataSetFromMatrix(
  countData=round(TE_counts[,rna_samples]),
  colData=TE_metadata[which(rownames(TE_metadata) %in% rna_samples), ],
  design=~Condition)
keep <- rowSums(counts(RNA_ddsMat) >= 10) >= 2
RNA_ddsMat <- RNA_ddsMat[keep,]
RNA_ddsMat <- DESeq(RNA_ddsMat)
RNA_ddsMat_Coll <- rlog( RNA_ddsMat, blind=FALSE)
RNA_ddsMat_Coll_mat <- assay(RNA_ddsMat_Coll)
RNA_ddsMat_Coll_cor <- cor(RNA_ddsMat_Coll_mat)

```

### RNA samples PCA.

```{r, echo=F, message=F, fig.width=10, fig.height=10}
vsd <- DESeq2::vst(RNA_ddsMat, blind = TRUE)
ntop = 500
Pvars <- rowVars(assay(vsd))
select <- order(Pvars, decreasing = TRUE)[seq_len(min(ntop, length(Pvars)))]
PCA <- prcomp(t(assay(vsd)[select, ]), scale = F)

#PCA <- prcomp(t(assay(vsd)), scale = F)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
dataGG = data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2], sampleName = row.names(colData(vsd)),colData(vsd))

p = qplot(PC1, PC2, data = dataGG, shape=Condition, size=I(3),  main = "PC1 vs PC2, top 500 variable genes") + labs(x = paste0("PC1, VarExp:", round(percentVar[1],4)), y = paste0("PC2, VarExp:", round(percentVar[2],4))) + theme_bw() + theme(legend.position="bottom")   +geom_label_repel(aes(label=row.names(dataGG)))
p + ggalt::geom_encircle(data=dataGG,aes(group=Condition, fill = Condition), s_shape=0.2, expand=0.07, alpha=0.2, show.legend = T) 


```


```{r, echo=F, message=F}
#Run deseq2
res_rna = results(RNA_ddsMat, name="Condition_Met_vs_Primary") 
res_rna = lfcShrink(RNA_ddsMat,res=res_rna, coef = "Condition_Met_vs_Primary",  type="apeglm")
subset_rna_metadata <- rna_metadata %>% dplyr::filter(Sample_name %in% rna_samples)
SQ_samples <- subset_rna_metadata$Sample_name[subset_rna_metadata$Condition=="Primary"]
print(paste0("printing Primary tumor samples ", SQ_samples))
ko_samples <- subset_rna_metadata$Sample_name[subset_rna_metadata$Condition=="Met"]
print(paste0("printing case metastasis samples ", ko_samples))
res_rna$RNA_SQ_mean_values <- rowMeans( counts(RNA_ddsMat,normalized=TRUE)[,SQ_samples])
res_rna$RNA_Met_mean_values <- rowMeans( counts(RNA_ddsMat,normalized=TRUE)[,ko_samples])

#Merge DESeq2 and annotation matrix
res_rna <- res_rna[order(res_rna[,6]),]
res_rna <- as.data.frame(res_rna)  
write.csv(res_rna, file="/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/DESeq2_RNA.csv", row.names = T, quote = F)
```


### Volcano plot Met vs Primary. 
```{r, echo=F, message=F, warning=F}
EnhancedVolcano(res_rna,lab = rownames(res_rna), x = 'log2FoldChange',y = 'pvalue', title = "Met vs Primary RNA results",  pCutoff = 1e-05, FCcutoff = 1)
```

### Heatmap showing expression of DE genes padj < 0.05 LFC 1
```{r, echo=F, message=F, warning=F}
#draw heatmap with the most disregulated genes
padj_genes <-na.omit(res_rna) %>% filter(padj<0.05 ,  log2FoldChange > 1 |  log2FoldChange < -1)

padj_genes2 <- rownames(padj_genes)
print(paste0(length(padj_genes2), " DE genes under p < 0.05 and LFC 1 in RNA_seq data"))

#Get the transformed values and filter genes
subset_mat <- RNA_ddsMat_Coll_mat[padj_genes2,]

#Now select the samples for each Condition
needed_samples <- c(SQ_samples,ko_samples)
subset_mat2 <- subset_mat[,needed_samples]
pheatmap(subset_mat2, scale="row", show_rownames = F, show_colnames = T, cluster_rows = T, legend = T, main =  "Heatmap of DE genes found in bulkRNAseq analysis", annotation_col = as.data.frame(colData(RNA_ddsMat))[, c("Data_type", "Condition")])
```


## {-}

## RIBOseq Met vs Primary DE results.{.tabset}


```{r, echo=F, message=F}
# RIBO seq data. PCA of the data.
#Create bulkRIBOseq object.
RIBO_samples <-TE_metadata[which(TE_metadata$Data_type == "RIBO_seq"),]$Sample_name 
RIBO_ddsMat = DESeqDataSetFromMatrix(
  countData=round(TE_counts[,RIBO_samples]),
  colData=TE_metadata[which(TE_metadata$Data_type == "RIBO_seq"),],
  design=~Condition)
keep <- rowSums(counts(RIBO_ddsMat) >= 10) >= 2
RIBO_ddsMat <- RIBO_ddsMat[keep,]
RIBO_ddsMat <- DESeq(RIBO_ddsMat)

RIBO_ddsMat_Coll <- rlog( RIBO_ddsMat, blind=FALSE)
RIBO_ddsMat_mat <- assay(RIBO_ddsMat_Coll)
```

### Sample distribution PCA

```{r, echo=F, message=F}
vsd <- DESeq2::vst(RIBO_ddsMat, blind = TRUE)
ntop = 500
Pvars <- rowVars(assay(vsd))
select <- order(Pvars, decreasing = TRUE)[seq_len(min(ntop, length(Pvars)))]
PCA <- prcomp(t(assay(vsd)[select, ]), scale = F)

#PCA <- prcomp(t(assay(vsd)), scale = F)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
dataGG = data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2], sampleName = row.names(colData(vsd)),colData(vsd))

p = qplot(PC1, PC2, data = dataGG, shape=Condition, size=I(3),  main = "PC1 vs PC2, top 500 variable genes") + labs(x = paste0("PC1, VarExp:", round(percentVar[1],4)), y = paste0("PC2, VarExp:", round(percentVar[2],4))) + theme_bw() + theme(legend.position="bottom")   +geom_label_repel(aes(label=row.names(dataGG)))
p + ggalt::geom_encircle(data=dataGG,aes(group=Condition, fill = Condition), s_shape=0.2, expand=0.07, alpha=0.2, show.legend = T) 
```

```{r, echo=F, message=F}
#Run deseq2
res_RIBO = results(RIBO_ddsMat, name="Condition_Met_vs_Primary") 
res_RIBO = lfcShrink(RIBO_ddsMat,res=res_RIBO, coef = "Condition_Met_vs_Primary",  type="apeglm")
SQ_samples <- ribo_metadata$Sample_name[ribo_metadata$Condition=="Primary"]
print(paste0("printing Primary tumor samples ", SQ_samples))
ko_samples <- ribo_metadata$Sample_name[ribo_metadata$Condition=="Met"]
print(paste0("printing case metastasis samples ", ko_samples))
res_RIBO$RIBO_SQ_mean_values <- rowMeans( counts(RIBO_ddsMat,normalized=TRUE)[,SQ_samples])
res_RIBO$RIBO_Met_mean_values <- rowMeans( counts(RIBO_ddsMat,normalized=TRUE)[,ko_samples])

#Merge DESeq2 and annotation matrix
res_RIBO <- res_RIBO[order(res_RIBO[,6]),]
res_RIBO <- as.data.frame(res_RIBO)  
write.csv(res_RIBO, file="/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/DESeq2_RIBO.csv", row.names = T, quote = F)
```



### Volcano plot Met vs Primary. 
```{r, echo=F, message=F, warning=F}
EnhancedVolcano(res_RIBO,lab = rownames(res_RIBO), x = 'log2FoldChange',y = 'pvalue', title = "Met vs Primary RIBO results",  pCutoff = 1e-05, FCcutoff = 1)
```

### Heatmap showing expression of DE genes padj < 0.05 LFC 1
```{r, echo=F, message=F, warning=F}
#draw heatmap with the most disregulated genes
padj_genes <-na.omit(res_RIBO) %>% filter(padj<0.05 ,  log2FoldChange > 1 |  log2FoldChange < -1)

padj_genes2 <- rownames(padj_genes)
print(paste0(length(padj_genes2), " DE genes under p < 0.05 and LFC 1 in RIBO_seq data"))

#Get the transformed values and filter genes
subset_mat <- RIBO_ddsMat_mat[padj_genes2,]

#Now select the samples for each Condition
needed_samples <- c(SQ_samples,ko_samples)
subset_mat2 <- subset_mat[,needed_samples]
pheatmap(subset_mat2, scale="row", show_rownames = F, show_colnames = T, cluster_rows = T, legend = T, main =  "Heatmap of DE genes found in RIBOseq analysis", annotation_col =as.data.frame(colData(RIBO_ddsMat))[, c("Data_type", "Condition")])
```


## {-}

## Plot TE results for each selenoprotein {.tabset} 

For each sample do log2(RIBO)-log(RNA) and plot the values. Across these values add the LFC of Met vs SQ obtained by DESeq2. Therefore: 

(log2(RIBO_met)-log(RNA_met)) - (log2(RIBO_sq)-log(RNA_sq))

Sample values should be in accordance with the values obtained by DESeq2.


To calculate TE for each sample, since we dont have coupled samples we use the following combinations since come from same mouse:
met_samples_2341 <- c("RIBO_seq_cm2341_met_liv", "RNA_seq_EP_6")
met_samples_2835 <- c("RIBO_seq_cm2835_met_liv", "RNA_seq_EP_19")

primary_2341 <- c("RIBO_seq_cm2341_sq", "RNA_seq_EP_2")
primary_2835 <- c("RIBO_seq_cm2835_sq", "RNA_seq_EP_16")


```{r, fig.height=10, fig.width=10}
TE_Xeno <- TE_normalized_counts
rownames(TE_Xeno) <- TE_Xeno$X
TE_Xeno$X<- NULL
remove_samples <- c("RNA_seq_EP_13", "RNA_seq_EP_11")
TE_Xeno <- TE_Xeno %>% select(!remove_samples)
met_samples_2341 <- c("RIBO_seq_cm2341_met_liv", "RNA_seq_EP_6")
met_samples_2835 <- c("RIBO_seq_cm2835_met_liv", "RNA_seq_EP_19")

primary_2341 <- c("RIBO_seq_cm2341_sq", "RNA_seq_EP_2")
primary_2835 <- c("RIBO_seq_cm2835_sq", "RNA_seq_EP_16")

met_2341_TE_Xeno <- TE_Xeno %>% select(all_of(met_samples_2341))
met_2835_TE_Xeno <- TE_Xeno %>% select(all_of(met_samples_2835))

primary_2341_TE_Xeno <- TE_Xeno %>% select(all_of(primary_2341))
primary_2835_TE_Xeno <- TE_Xeno %>% select(all_of(primary_2835))

TE_values <- function(dataframe, sample_name){
  TE_values_primary <- data.frame()
  for (gene in seleno_list){
    dataframe2 <- data.frame(gene=gene)
    sub_test <- dataframe[gene,]
    res <- as.data.frame(log2(sub_test[,grep("RIBO", colnames(sub_test), value = T)])- log2(sub_test[,grep("RNA", colnames(sub_test), value = T)]))
    res <- res %>% dplyr::rename(!!paste0(sample_name):=1)
    dataframe2 <- cbind(dataframe2, res)
    TE_values_primary <- rbind(TE_values_primary,dataframe2)
  }
  TE_values_primary <- na.omit(TE_values_primary)
  return(TE_values_primary)
}



primary_2341_TE_Xeno_values <- TE_values(primary_2341_TE_Xeno, "primary_2341")
primary_2835_TE_Xeno_values <- TE_values(primary_2835_TE_Xeno, "primary_2835")
met_2341_TE_Xeno_values <- TE_values(met_2341_TE_Xeno, "met_2341")
met_2835_TE_Xeno_values <- TE_values(met_2835_TE_Xeno, "met_2835")


xeno_list <- list(met_2341_TE_Xeno_values,met_2835_TE_Xeno_values,primary_2341_TE_Xeno_values, primary_2835_TE_Xeno_values)
Xeno_TE_table_by_sample <- join_all(xeno_list, type="inner")

Xeno_TE_table_by_sample[Xeno_TE_table_by_sample=="Inf"] <- 0

#Add DESeq2 results.
TE_Xeno <- results_TE
subset_TE_Xeno <- TE_Xeno %>% dplyr::rename(LFC_Met_Prim =log2FoldChange, gene=X) %>% dplyr::filter(gene %in% seleno_list) %>% dplyr::select(gene, LFC_Met_Prim)
Xeno_TE_table_by_sample <- join(Xeno_TE_table_by_sample,subset_TE_Xeno, type="inner") 

Xeno_TE_table_by_sample <- Xeno_TE_table_by_sample %>% arrange(LFC_Met_Prim)
rownames(Xeno_TE_table_by_sample) <- Xeno_TE_table_by_sample$gene
annotation_row = data.frame(Xeno_TE_table_by_sample$LFC_Met_Prim)
rownames(annotation_row) = Xeno_TE_table_by_sample$gene
names(annotation_row) <- "TE_LFC"
Xeno_TE_table_by_sample$gene <- NULL
write.csv(Xeno_TE_table_by_sample, "/SSD/maa7095/RIBOseq/Shino_new_analysis/inframe_downtream_analysis/Xenograft/TE_analysis/table_for_heatmaps/Xenograft_TE_table.csv")

condition <- data.frame(Condition=c("Met", "Met", "Primary", "Primary"))

Xeno_TE_table_by_sample <- as.matrix(Xeno_TE_table_by_sample[,1:4])
order_samples <- colnames(Xeno_TE_table_by_sample)
paletteLength <- 50
myColor <- colorRampPalette(c("navy", "white", "red"))(paletteLength)

myBreaks <- c(seq(min(Xeno_TE_table_by_sample), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(Xeno_TE_table_by_sample)/paletteLength, max(Xeno_TE_table_by_sample), length.out=floor(paletteLength/2)))

annotation_colors = list(
  Condition=c(Met="salmon1", Primary="slategray"),
  TE_LCF = colorRampPalette("purple"))



pheatmap(Xeno_TE_table_by_sample, cluster_col = F, legend = T, main =  "Translation Efficiency in Xenograft data", breaks = myBreaks, color = myColor, annotation_col = as.data.frame(condition), annotation_colors = annotation_colors, cellwidth = 25, cellheight = 25, annotation_row =as.data.frame(annotation_row), border_color=NA, treeheight_row=0,treeheight_col=0 , cluster_rows = F, name="TE")
```

## {-}





# Session info.
```{r, echo=F, message=F}
sessionInfo()
```

